---
title: "DM Group 19"
author: "Yang Le Lim"
date: "4/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# To filter compas two years

Subset of columns used in original analysis are:
raw_data, age, c_charge_degree, race, age_cat, score_text, sex, priors_count, days_b_screening_arrest, decile_score, is_recid, two_year_recid, c_jail_in, c_jail_out

```{r}
raw_data <- read.csv("./compas-scores-two-years.csv")
nrow(raw_data)

compas_non_violent <- raw_data %>% 
  filter(days_b_screening_arrest <= 30) %>%
  filter(days_b_screening_arrest >= -30) %>%
  filter(is_recid != -1) %>%
  filter(c_charge_degree != "O") %>%
  filter(score_text != 'N/A')
nrow(compas_non_violent)
```

# To filter compas two years violent 

Subset of columns used in original analysis are:
raw_data, age, c_charge_degree, race, age_cat, v_score_text, sex, priors_count, 
days_b_screening_arrest, v_decile_score, is_recid, two_year_recid

```{r}
raw_data.2 <- read.csv("./compas-scores-two-years-violent.csv")
nrow(raw_data.2)

compas_violent <- raw_data.2 %>%
        filter(days_b_screening_arrest <= 30) %>%
        filter(days_b_screening_arrest >= -30) %>% 
        filter(is_recid != -1) %>%
        filter(c_charge_degree != "O") %>%
        filter(v_score_text != 'N/A')
nrow(compas_violent)

```