---
title: "DM Group 19 Decision Tree and Random Forest Analysis"
author: "Samvrudhi Shankar"
date: "5/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rpart)
library(randomForest)
library(tidyverse)
library(ggplot2)
library(ISLR)
library(MASS)
library(klaR) 
library(knitr)
library(glmnet)
library(gam)
library(plyr)
library(reshape)
library(boot)
library(survival)
library(ggfortify)
library(gridExtra)
library(grid)
```

Summary:

We seek to explore predictors of recidivism on the Broward County Population by conducting the following analysis:
(1) Construct an RAI for predicting two-year recidivism on the Broward County population. 
(2) Construct an RAI for predicting two-year violent recidivism on the Broward County population
(3) Compare the performance of our RAIs to COMPAS
(4) Determine whether each RAI is equally predictive across race, age, and gender

This file creates four models: (1) Decision tree to classify people who have committed a criminal offense in Broward County as likely to recidivate or not,  (2) Decision tree to classify people who have committed a criminal offense in Broward County as likely to violently recidivate or not, (3) Random forest to classify people who have committed a criminal offense in Broward County as likely to recidivate or not, and (4) Random forest to classify people who have committed a criminal offense in Broward County as likely to violently recidivate or not. These four models encompass steps 1 and 2 of our analysis above.

The nonviolent decision tree model has an accuracy of approximately 68%, whereas the violent decision tree model has an accuracy of about 89%. The nonviolent random forest model has an accuracy of approximately 68%, whereas the violent random forest model has an accuracy of about 89%.

The nonviolent decision tree model had a concordance of 61.2% and the nonviolent random forest model had a concordance of 62.8%, whereas COMPAS had a concordance of 63.6% for nonviolent recidivism. The violent random forest model had a concordance of 50.2%, whereas COMPAS had a concordance of 65.1% for violent recidivism. We were unable to determine the concordance for the violent decision tree model because it predicted that all of the observations were 0, or at low risk for violently recidivating. The cox proportional hazard model that we built to calculate concordance encompasses step 3 of our analysis above. 

Because the violent decision tree model was unable to distinguish between high and low risk defendants, we chose to use random forests only in step 4 of our analysis, where we determined whether random forests are equally predictive across race, age, and gender for nonviolent and violent recidivism. We found that assigning a 1, or a "high risk" label to a white defendant, is more indicative of their likelihood of nonviolently recidivating than assigning a 1 to a black defendant. Assigning a 1, or a "high risk" label to a person over age 45 is more indicative of their likelihood of nonviolently recidivating than assigning a 1 to a person under age 25. Also, assigning a 1, or a "high risk" label to a male defendant is more indicative of their likelihood of nonviolently recidivating than assigning a 1 to a female defendant.

We found that assigning a 1, or a "high risk" label to a black defendant is more indicative of their likelihood of violently recidivating than assigning a 1 to a white defendant. Assigning a 1, or a "high risk" label to a defendant under age 25 is more indicative of their likelihood of violently recidivating than assigning a 1 to a defendant over age 45. In addition, assigning a 1 or a "high risk" label to a female defendant is more indicative of their likelihood of violently recidivating than assigning a 1 to a male defendant. Therefore, the nonviolent and violent random forest models are NOT equally predictive across race, age, or gender.


```{r}
# Read in compas data
compas_non_violent <- read.csv("./compas-scores-two-years.csv")
compas_non_violent$is_recid <- factor(compas_non_violent$is_recid)
compas_non_violent$is_violent_recid <- factor(compas_non_violent$is_violent_recid)
index = sample( 1:nrow( compas_non_violent ), round( nrow( compas_non_violent )*0.6 ), replace = FALSE )
train = compas_non_violent[ index, ] # About 60% of the observations
test = compas_non_violent[ -index, ] # About 40% of the observations

# Read in cox proportional hazards data
cox_parsed <- read.csv("./cox-parsed.csv")

```

# Variable Initialization
```{r}

required_columns <- c("sex",
                      "age",
                      "race",
                      "juv_fel_count",
                      "juv_misd_count",
                      "juv_other_count",
                      "priors_count")
dependent_column_non_violent <- "is_recid"
formula_non_violent <- paste0(dependent_column_non_violent,
                              "~",
                              paste0(required_columns, collapse = "+"))

dependent_column_violent <- "is_violent_recid"
formula_violent <- paste0(dependent_column_violent,
                              "~",
                              paste0(required_columns, collapse = "+"))
```

# Decision Tree Analysis
```{r}
#-----------------------------------------------NONVIOLENT DECISION TREE-----------------------------------------------------------------

dt_object <- rpart(formula_non_violent, data=train)
dt_value = predict(object=dt_object,
                   newdata =test[required_columns],
                   type="class" )

dt_error = sum(test[[dependent_column_non_violent]] != dt_value)  / length(dt_value)
dt_error

# Create confusion matrix
tab <- table(dt_value, test$is_recid)
tab

# Test Accuracy: 1960/2886 = 67.914% (exact numbers will change after every run, but accuracy and error rate will remain approximately the same)
# Test error rate: 32.086%


#-----------------------------------------------VIOLENT DECISION TREE------------------------------------------------------------------------

dt_object_v <- rpart(formula_violent, data=train)
dt_value_v = predict(object=dt_object_v,
                   newdata =test[required_columns],
                   type="class" )
dt_error_v = sum(test[[dependent_column_violent]] != dt_value_v)  / length(dt_value_v)
dt_error_v

# Create confusion matrix
tab <- table(observed = test$is_violent_recid, predicted = dt_value_v)
tab

# Test Accuracy: 89.085%
# Test Error Rate: 10.915%


```


# Cox Proportional Hazards Decision Tree
```{r}

# Run a Cox Proportional Hazards model to compare the accuracy of the decision trees to COMPAS' model

# Create new dataframe test_with_labels that appends labels to test dataframe
test_with_labels <- test
test_with_labels$nonviolent_dt_labels = dt_value
test_with_labels$violent_dt_labels = dt_value_v

# Write decision tree predictions to cox-parsed-dt.csv 
cox_parsed_dt = cox_parsed[FALSE,]
for (i in (1: nrow(test_with_labels))) {
  new_rows <- subset(cox_parsed, name == as.character(test_with_labels[i,]$name))
  new_rows$non_violent_dt_labels = test_with_labels[i, ]$nonviolent_dt_labels
  new_rows$violent_dt_labels = test_with_labels[i, ]$violent_dt_labels
  cox_parsed_dt <- rbind(cox_parsed_dt, new_rows)
}


#write.csv(cox_parsed_dt, "cox-parsed-dt.csv")

data <- filter(filter(read.csv("./cox-parsed-dt.csv"), score_text != "N/A"), end > start) %>%
        mutate(race_factor = factor(race,
                                  labels = c("African-American", 
                                             "Asian",
                                             "Caucasian", 
                                             "Hispanic", 
                                             "Native American",
                                             "Other"))) %>%
        within(race_factor <- relevel(race_factor, ref = 3)) %>%
        mutate(score_factor = factor(score_text)) %>%
        within(score_factor <- relevel(score_factor, ref=2)) %>%
        mutate(non_violent_score_factor = factor(non_violent_dt_labels)) %>%
        within(non_violent_score_factor <- relevel(non_violent_score_factor, ref = 1)) %>%
        mutate(violent_score_factor = factor(violent_dt_labels)) %>%
        within(violent_score_factor <- relevel(violent_score_factor, ref = 1))
    

f <- Surv(start, end, event, type="counting") ~ non_violent_score_factor
nonviolent_model <- coxph(f, data=data)
summary(nonviolent_model)
# Nonviolent decision tree model has a concordance of 61.2%
# COMPAS system's concordance is 63.6%



f <- Surv(start, end, event, type="counting") ~ violent_score_factor
#violent_model <- coxph(f, data=data)
#summary(violent_model)
# Unable to calculate concordance of violent decision tree model 


```


# Random Forest Analysis
```{r}

# -------------------------------------------------------------NONVIOLENT RANDOM FOREST-----------------------------------------------------------
rf_object <- randomForest(y=train[[dependent_column_non_violent]],
                          x=train[required_columns])

rf_value = predict(object=rf_object,
                   newdata =test[required_columns],
                   type="class" )
rf_error = sum(test[[dependent_column_non_violent]] != rf_value)  / length(rf_value)
rf_error

# Create confusion matrix
tab <- table(rf_value, test$is_recid)
tab

# Test Accuracy: 68.157%
# Test Error Rate: 31.982%


# -----------------------------------------------------------------VIOLENT RANDOM FOREST-------------------------------------------------------------
rf_object_v <- randomForest(y=train[[dependent_column_violent]],
                          x=train[required_columns])

rf_value_v = predict(object=rf_object_v,
                   newdata =test[required_columns],
                   type="class" )
rf_error_v = sum(test[[dependent_column_violent]] != rf_value_v)  / length(rf_value_v)
rf_error_v

# Create confusion matrix
tab <- table(rf_value_v, test$is_violent_recid)
tab

# Test Accuracy: 89.051%
# Test Error Rate: 10.949%

```


# Cox Proportional Hazards Random Forest
```{r}

# Run a Cox Proportional Hazards model to compare the accuracy of the random forests to COMPAS' model

# Create new dataframe test_with_labels that appends labels to test dataframe
test_with_labels <- test
test_with_labels$nonviolent_rf_labels = rf_value
test_with_labels$violent_rf_labels = rf_value_v

# Write decision tree predictions to cox-parsed-rf.csv 
cox_parsed_rf = cox_parsed[FALSE,]
for (i in (1: nrow(test_with_labels))) {
  new_rows <- subset(cox_parsed, name == as.character(test_with_labels[i,]$name))
  new_rows$non_violent_rf_labels = test_with_labels[i, ]$nonviolent_rf_labels
  new_rows$violent_rf_labels = test_with_labels[i, ]$violent_rf_labels
  cox_parsed_rf <- rbind(cox_parsed_rf, new_rows)
}


#write.csv(cox_parsed_rf, "cox-parsed-rf.csv")

data <- filter(filter(read.csv("./cox-parsed-rf.csv"), score_text != "N/A"), end > start) %>%
        mutate(race_factor = factor(race,
                                  labels = c("African-American", 
                                             "Asian",
                                             "Caucasian", 
                                             "Hispanic", 
                                             "Native American",
                                             "Other"))) %>%
        within(race_factor <- relevel(race_factor, ref = 3)) %>%
        mutate(score_factor = factor(score_text)) %>%
        within(score_factor <- relevel(score_factor, ref=2)) %>%
        mutate(non_violent_score_factor = factor(non_violent_rf_labels)) %>%
        within(non_violent_score_factor <- relevel(non_violent_score_factor, ref = 1)) %>%
        mutate(violent_score_factor = factor(violent_rf_labels)) %>%
        within(violent_score_factor <- relevel(violent_score_factor, ref = 1))  %>%
        mutate(age_factor = factor(age_cat))  %>%
        within(age_factor <- relevel(age_factor, ref = 3))


white <- filter(data, race == "Caucasian")
white_fit <- survfit(f, data=white)
black <- filter(data, race == "African-American")
black_fit <- survfit(f, data=black)
    

f_nonviolent <- Surv(start, end, event, type="counting") ~ non_violent_score_factor
nonviolent_model <- coxph(f_nonviolent, data=data)
summary(nonviolent_model)
# Nonviolent random forest model has a concordance of 62.8%
# COMPAS system's concordance is 63.6%

# Survival Plot Nonviolent Random Forest
fit <- survfit(f_nonviolent, data=data)
plotty <- function(fit, title) {
  return(autoplot(fit, conf.int=T, censor=F) + ggtitle(title) + ylim(0,1))
}
plotty(fit, "Overall Nonviolent")

# People classified as high risk do recidivate at higher rates for nonviolent data.


f_violent <- Surv(start, end, event, type="counting") ~ violent_score_factor
violent_model <- coxph(f_violent, data=data)
summary(violent_model)
# Violent random forest model has a concordance of 50.2%
# COMPAS's violent recidivism score has a concordance score of 65.1%


# Survival Plot Violent Random Forest
fit <- survfit(f_violent, data=data)
plotty <- function(fit, title) {
  return(autoplot(fit, conf.int=T, censor=F) + ggtitle(title) + ylim(0,1))
}
plotty(fit, "Overall Violent")

# People classified as high risk do recidivate at higher rates for violent data, but the confidence band indicates that this may not always be the case.


```

# Determine whether random forests are equally predictive across race, age, and gender for nonviolent recidivism
```{r}

hist(compas_non_violent$age)
table(compas_non_violent$age_cat)
table(compas_non_violent$race)
table(compas_non_violent$gender)

# ---------------------------------------------------------NONVIOLENT-------------------------------------------------------------------
# Race
f2 <- Surv(start, end, event, type="counting") ~ race_factor + non_violent_score_factor + race_factor * non_violent_score_factor
model2 <- coxph(f2, data=data)
print(summary(model2))

# High-risk black defendants (assigned a label of 1) are approx. 2.6 times more likely to nonviolently recidivate than low-risk black defendants (assigned a label of 0).
print(paste("Black High Hazard:", exp(-.1601 + 1.126), sep=" "))

# High-risk white defendants (assigned a label of 1) are approx. 3.1 times more likely to nonviolently recidivate than low-risk white defendants (assigned a label of 0).
print(paste("White High Hazard:", exp(1.126), sep=" "))

# Assigning a 1, or a "high risk" label to a white defendant is more indicative of their likelihood of nonviolently recidivating than assigning a 1 to a black defendant.

# Race specific cox models have somewhat similar concordance values, but the concordance for black defendants is lower.
summary(coxph(f_nonviolent, data=white)) # Concordance of 62.7% 
summary(coxph(f_nonviolent, data=black)) # Concordance of 61.1%

------------------------------------------------------------------------------------------------------------------------------------------
# Age
f3 <- Surv(start, end, event, type="counting") ~ age_factor + non_violent_score_factor + age_factor * non_violent_score_factor
model3 <- coxph(f3, data=data)
print(summary(model3))

# High-risk defendants over age 45 are approx. 3.3 times more likely to nonviolently recidivate than low-risk defendants over age 45.
print(paste("Over Age 45 High Hazard:", exp(0.3026 + 0.8907), sep=" "))

# High-risk defendants under age 25 are approx. 2.4 times more likely to nonviolently recidivate than low-risk defendants under age 25. 
print(paste("Under Age 25 High Hazard:", exp(0.8907), sep=" "))

# Assigning a 1, or a "high risk" label to a person over age 45 is more indicative of their likelihood of nonviolently recidivating than assigning a 1 to a person under age 25. 
-------------------------------------------------------------------------------------------------------------------------------------------
# Gender
f4 <- Surv(start, end, event, type="counting") ~ sex + non_violent_score_factor + sex * non_violent_score_factor
model4 <- coxph(f4, data=data)
print(summary(model4))

# High-risk male defendants are approx. 2.9 times more likely to nonviolently recidivate than low-risk male defendants.
print(paste("Male High Hazard:", exp(0.0952 + 0.9696), sep=" "))

# High-risk female defendants are approx. 2.6 times more likely to nonviolently recidivate than low-risk female defendants.
print(paste("Female High Hazard:", exp(0.9696), sep=" "))

# Assigning a 1, or a "high risk" label to a male defendant is more indicative of their likelihood of nonviolently recidivating than assigning a 1 to a female defendant.

# Survival Plot Females and Males
female <- filter(data, sex == "Female")
male   <- filter(data, sex == "Male")
male_fit <- survfit(f_nonviolent, data=male)
female_fit <- survfit(f_nonviolent, data=female)
grid.arrange(plotty(female_fit, "Female"), plotty(male_fit, "Male"),ncol=2)
```


# Determine whether random forests are equally predictive across race, age, and gender for violent recidivism
```{r}

# ------------------------------------------------------------VIOLENT-----------------------------------------------------------------------
# Race

f5 <- Surv(start, end, event, type="counting") ~ race_factor + violent_score_factor + race_factor * violent_score_factor
model5 <- coxph(f5, data=data)
print(summary(model5))

# High-risk black defendants (assigned a label of 1) are approx. 3.4 times more likely to violently recidivate than low-risk black defendants (assigned a label of 0).
print(paste("Black High Hazard:", exp(0.17908 + 1.05811), sep=" "))

# High-risk white defendants (assigned a label of 1) are approx. 2.9 times more likely to violently recidivate than low-risk white defendants (assigned a label of 0).
print(paste("White High Hazard:", exp(1.05811), sep=" "))

# Assigning a 1, or a "high risk" label to a black defendant is more indicative of their likelihood of violently recidivating than assigning a 1 to a white defendant.

# Race specific cox models have very similar concordance values
summary(coxph(f_violent, data=white)) # Concordance of 50.1% 
summary(coxph(f_violent, data=black)) # Concordance of 50.3%

----------------------------------------------------------------------------------------------------------------------------------------------
# Age
  
f6 <- Surv(start, end, event, type="counting") ~ age_factor + violent_score_factor + age_factor * violent_score_factor
model6 <- coxph(f6, data=data)
print(summary(model6))

# High-risk defendants over age 45 are approx. 2.9 times more likely to violently recidivate than low-risk defendants over age 45.
print(paste("Over Age 45 High Hazard:", exp(-0.07077 + 1.13375), sep=" "))

# High-risk defendants under age 25 are approx. 3.1 times more likely to violently recidivate than low-risk defendants under age 25. 
print(paste("Under Age 25 High Hazard:", exp(1.13375), sep=" "))

# Assigning a 1, or a "high risk" label to a defendant under age 25 is more indicative of their likelihood of violently recidivating than assigning a 1 to a defendant over age 45.
------------------------------------------------------------------------------------------------------------------------------------------------
# Gender
  
f7 <- Surv(start, end, event, type="counting") ~ sex + violent_score_factor + sex * violent_score_factor
model7 <- coxph(f7, data=data)
print(summary(model7))

# High-risk male defendants are approx. 3.1 times more likely to violently recidivate than low-risk male defendants.
print(paste("Male High Hazard:", exp(-0.80590 + 1.94013), sep=" "))

# High-risk female defendants are approx. 7.0 times more likely to violently recidivate than low-risk female defendants.
print(paste("Female High Hazard:", exp(1.94013), sep=" "))

# Assigning a 1, or a "high risk" label to a female defendant is more indicative of their likelihood of violently recidivating than assigning a 1 to a male defendant.

# Survival Plot Females and Males
male_fit <- survfit(f_violent, data=male)
female_fit <- survfit(f_violent, data=female)
grid.arrange(plotty(female_fit, "Female"), plotty(male_fit, "Male"),ncol=2)
```



